CUNY SPS DATA 605 Final

### 2019-05-18 DATA605/MSilva_Final.Rmd
Work on DATA605 Final
More work on 605 final

---
title: "DATA 605 Final"
author: "Mike Silva"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

```{r, comment=NA}
set.seed(12345)
N <- 10
n <- 10000
mu <- sigma <- (N + 1)/2
df <- data.frame(X = runif(n, min=1, max=N), 
                 Y = rnorm(n, mean=mu, sd=sigma))
summary(df$X)
summary(df$Y)
```

*Probability.* Calculate as a minimum the below probabilities a through c. Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable. Interpret the meaning of all probabilities.

```{r}
x <- median(df$X)
y <- as.numeric(quantile(df$Y)["25%"])
```

@@

```{r}
P_a_and_b <- df %>%
  filter(X > x,
         X > y) %>%
  nrow() / n

@@

```{r}
problem_1b <- df %>%
  filter(X > x,
         Y > y) %>%
  nrow() / n
```

@@

```{r}
P_a_and_b <- df %>%
  filter(X < x,
         X > y) %>%
  nrow() / n

@@

```{r}
# Create Joint Probabilities
temp <- df %>%
  mutate(A = ifelse(X > x, " X greater than x", " X not greater than x"),
         B = ifelse(Y > y, " Y greater than y", " Y not greater than y")) %>%
  group_by(A, B) %>%
  summarise(count = n()) %>%
  mutate(probability = count / n)

@@

```{r, comment=NA}
count_data <- temp %>%
  filter(A != "Total",
         B != "Total") %>%
  select(-probability) %>%
  spread(A, count) %>%
  as.data.frame()

@@

```{r, comment=NA}
chisq.test(count_data)
```

@@

*Descriptive and Inferential Statistics.* Provide univariate descriptive statistics and appropriate plots for the training data set. Provide a scatterplot matrix for at least two of the independent variables and the dependent variable. Derive a correlation matrix for any three quantitative variables in the dataset. Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval. Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?

*Linear Algebra and Correlation.* Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.

*Calculus-Based Probability & Statistics.* Many times, it makes sense to fit a closed form distribution to data. Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary. Then load the MASS package and run
fitdistr to fit an exponential probability density function. (See
https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html). Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)). Plot a histogram and compare it with a histogram of your original variable. Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF). Also generate a 95% confidence interval from the empirical data, assuming normality. Finally, provide the empirical 5th percentile and 95th percentile of the data. Discuss.

*Modeling.* Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.

>>>>

---
title: "CUNY SPS DATA 605 Final"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#output: html_document
#output:
#  rmdformats::html_clean:
#    highlight: kate
library(knitr)
#library(rmdformats)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

```{r, echo=TRUE, eval=TRUE, comment=NA}
set.seed(12345)
N <- 10
n <- 10000
mu <- sigma <- (N + 1)/2
df <- data.frame(X = runif(n, min=1, max=N), 
                 Y = rnorm(n, mean=mu, sd=sigma))
summary(df$X)
summary(df$Y)
```

### Probability

Calculate as a minimum the below probabilities a through c. Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable. Interpret the meaning of all probabilities.

```{r echo=TRUE, eval=TRUE}
x <- median(df$X)
y <- as.numeric(quantile(df$Y)["25%"])
```

@@

```{r echo=TRUE, eval=TRUE}
P_a_and_b <- df %>%
  filter(X > x,
         X > y) %>%
  nrow() / n

@@

```{r echo=TRUE, eval=TRUE}
problem_1b <- df %>%
  filter(X > x,
         Y > y) %>%
  nrow() / n
```

@@

```{r echo=TRUE, eval=TRUE}
P_a_and_b <- df %>%
  filter(X < x,
         X > y) %>%
  nrow() / n

@@

```{r echo=TRUE, eval=TRUE}
# Create Joint Probabilities
temp <- df %>%
  mutate(A = ifelse(X > x, " X greater than x", " X not greater than x"),
         B = ifelse(Y > y, " Y greater than y", " Y not greater than y")) %>%
  group_by(A, B) %>%
  summarise(count = n()) %>%
  mutate(probability = count / n)

@@

```{r, echo=TRUE, eval=TRUE, comment=NA}
count_data <- temp %>%
  filter(A != "Total",
         B != "Total") %>%
  select(-probability) %>%
  spread(A, count) %>%
  as.data.frame()

@@

```{r, echo=TRUE, eval=TRUE, comment=NA}
chisq.test(count_data)
```

@@

```{r, echo=TRUE, eval=TRUE}
kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
  mutate(MSSubClass = as.factor(MSSubClass))
```

### Descriptive and Inferential Statistics

Provide univariate descriptive statistics and appropriate plots for the training data set.

```{r echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
for (variable in  names(kaggle)){
  cat(paste("<h4>", variable, "</h4>"))

  d <- kaggle[,names(kaggle) == variable]

  if(is.factor(d)){
    p <- ggplot(kaggle, aes_string(variable)) + 
      geom_bar() + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

    d_table <- kaggle %>%
      group_by(.dots = variable) %>%
      summarise(count = n()) %>%
      kable() %>%
      kable_styling()
  } else {
    p <- ggplot(kaggle, aes_string(variable)) + 
      geom_histogram(binwidth = 10)
    q <- quantile(d, na.rm = TRUE)
    d_table <- data.frame(Min = min(d, na.rm = TRUE),
                          First = q[2],
                          Median = median(d, na.rm = TRUE), 
                          Mean = mean(d, na.rm = TRUE),
                          Third = q[4],
                          Max = max(d, na.rm = TRUE)) %>%
      rename("First Q" = First,
             "Third Q" = Third) %>%
      gather(Measure, Value) %>%
      kable() %>%
      kable_styling()
  }
  print(p)
  print(d_table)
}
```

Provide a scatterplot matrix for at least two of the independent variables and the dependent variable.

```{r, echo=TRUE}
kaggle %>%
  select(TotalBsmtSF, GrLivArea, YearBuilt, SalePrice) %>%
  pairs()
```

Derive a correlation matrix for any three quantitative variables in the dataset.

```{r, echo=TRUE}
kaggle %>%
  select(TotalBsmtSF, GrLivArea, YearBuilt) %>%
  cor() %>%
  kable() %>%
  kable_styling()
```

Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval. Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?

```{r}
d_table <- data.frame(Min = min(d, na.rm = TRUE),
                          "First Qu." = q[2],
                          Median = median(d, na.rm = TRUE), 
                          Mean = mean(d, na.rm = TRUE),
                          "Third Qu." = q[4],
                          Max = max(d, na.rm = TRUE)) %>%
  gather(Measure, Value)

```

### Linear Algebra and Correlation

Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.

### Calculus-Based Probability & Statistics

Many times, it makes sense to fit a closed form distribution to data. Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary. Then load the MASS package and run fitdistr to fit an exponential probability density function. (See https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html). Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)). Plot a histogram and compare it with a histogram of your original variable. Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF). Also generate a 95% confidence interval from the empirical data, assuming normality. Finally, provide the empirical 5th percentile and 95th percentile of the data. Discuss.

### Modeling

Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.

(http://jse.amstat.org/v19n3/decock.pdf)
